---
title: "p8105_hw6_xs2330"
author: "Xiao Shi"
date: "November 15, 2018"
output: github_document
---
## Problem 1
Using tidyverse library for data cleaning and analysis, leaps for stepwise selection, and the rest for cross validation
```{r setup, warning= FALSE, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(leaps)
library(mgcv)
```

**Data cleaning and tidying**
```{r, message = FALSE, warning= FALSE}
raw_homicide = read_csv("./homicide-data.csv") 
homicide = raw_homicide %>%
  mutate(city_state = str_c(city, state, sep = ", ")) %>%
  mutate(case_status = recode(disposition, "Closed without arrest" = "unsolved", 
                      "Open/No arrest" = "unsolved",
                      "Closed by arrest" = "solved")) %>%
  filter(city_state != "Dallas, TX" & city_state != "Phoenix, AZ" & city_state != "Kansas City, MO" & city_state != "Tulsa, AL") %>%
  mutate(race_general = ifelse(victim_race == "White", "white","non_white"),
         victim_age = as.numeric(victim_age),
         race_general = fct_relevel(as.factor(race_general),"white")) %>%
  janitor::clean_names()

```
Upon removing cities that do not have information that we need, and mutate variables such as `victim_age` and `victim_race`, the resulting dataset consists `r nrow(homicide)` rows by `r ncol(homicide)` columns.

**Fitting model for Baltimore, MD**

This is a logistic regression with resolved vs unresolved as the outcome and `victim_age`, `victim_sex` and `race_general` (as just defined) as predictors.
```{r}
balti_homicide = homicide %>%
  filter(city_state == "Baltimore, MD") %>%
  mutate(case_status = ifelse(case_status == "solved",1,0))

glm_case_status = glm(case_status ~ victim_age + victim_sex + race_general, data = balti_homicide, family = binomial())
output_glm_case_status = summary(glm_case_status)
save(output_glm_case_status,file = "./homicide_balti.rdata")
```

Tidy the results and compare the adjusted odds ratio for solving homicides comparing white victims to non-white victims keeping all other variables fixed in Baltimore, MD.
```{r}
glm_case_status %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         OR_lower = exp(estimate - std.error * 1.96),
         OR_upper = exp(estimate + std.error * 1.96)) %>%
  select(term, OR, OR_lower, OR_upper, p.value) %>% 
  knitr::kable(digits = 3)
```

From the above table, we see that homicide cases among non-whites are much less likely to be solved compared to cases among white people. The adjusted odds ratio is 0.441 with a p-value of 0.0000027, indicating statistically significant difference at a 95% confidence level.

